Underdispersion in the context of community athletics

### 2024-03-28 inst/parkrun_idea.Rmd
tweaks

f <- function(n,alpha){
  jj <- alpha^(1:n)
  return(jj/sum(jj))
}

@@

```{r calculateLikemat,cache=FALSE}
alpha_try <- seq(from=1,to=6,len=10)
beta_try <- 1:26
calculate_from_scratch <- FALSE
if(calculate_from_scratch){
    M <- as.matrix(expand.grid(alpha=alpha_try, beta=beta_try))
    system.time(L <- t(apply(M,1,getlike)))
    colnames(L) <- c(letters,'alpha','beta')
    write.table(L,file="likelihoodmatrix.txt")
} else {
    L <- read.table("likelihoodmatrix.txt")
}
L_integer <- L  # counts
L[,1:26] <- L[,1:26]/mean(rowSums(L[,1:26]))
head(L_integer)
```

@@

## Now 10 in-silico races:
```{r doinsilico,cache=FALSE}
alphatrue <- 1.3
betatrue <- 20   # 13th letter = "m"
set.seed(1)
D <- replicate(10,which(rrace(setNames(f(26,alphatrue),letters))==letters[betatrue]))
sort(D)
D <- 10:20
D
```

>>>>

f <- function(n,alpha){
  jj <- 1/alpha^(1:n)
  return(jj/sum(jj))
}

@@

```{r calculateLikemat,cache=FALSE}
alpha_try <- seq(from=1,to=6,len=20)
beta_try <- 1:26
calculate_from_scratch <- FALSE
if(calculate_from_scratch){
    M <- as.matrix(expand.grid(alpha=alpha_try, beta=beta_try))
    system.time(L <- t(apply(M,1,getlike)))
    colnames(L) <- c(letters,'alpha','beta')
    write.table(L,file="likelihoodmatrix.txt")
} else {
    L <- read.table("likelihoodmatrix.txt")
}
L_integer <- L  # counts
L[,1:26] <- L[,1:26]/mean(rowSums(L[,1:26]))
head(L_integer)
```

@@

## Now 10 in-silico races:
```{r doinsilico,cache=FALSE}
alphatrue <- 2.3
betatrue <- 4  
set.seed(1)
D <- replicate(10,which(rrace(setNames(f(26,alphatrue),letters))==letters[betatrue]))
sort(D)
```
